---
title: "Class 12: Introduction to Genome Informatics Lab"
author: "Yu (Ericsson) Cao (PID: A16421048)"
date: "02-16-2024"
format: pdf
---





# Section 1. Proportion of G/G in a population

Downloaded a CSV file from Ensemble:  <https://useast.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39894595-39895595;v=rs8067378;vdb=variation;vf=959672880#373531_tablePanel>

Here we read this CSV file

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```

```{r}
table(mxl$Genotype..forward.strand.)
```


```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl) * 100
```

Now let's look at a different population. We picked the GBR.

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(gbr)
```

Find proportion of G|G

```{r}
round(table(gbr$Genotype..forward.strand.) / nrow(gbr) *100, 2)
```

This variant that is associated with childhood asthma is more frequent in the GBR population than the MXL population.

Let's now dig into this further.

## Section 4: Population Scale Analysis
One sample is obviously not enough to know what is happening in a population. You are interested in assessing genetic differences on a population scale. 

How many samples do we have?

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

```{r}
nrow(expr)
```

```{r}
table(expr$geno)
```

```{r}
library(ggplot2)

```

Let's make a boxplot

```{r}
library(plyr)

p_meds <- ddply(expr, .(geno), summarise, med = median(exp)) #median displayed

ggplot(expr) + aes(geno, exp, fill=geno) +
    xlab("Genotype") + 
    ylab("Expression") +
    ggtitle("Population Scale Analysis of ORMDL3") +
  geom_boxplot(notch=T) +
   geom_jitter(color="black", size=0.4, alpha=0.2) +
  geom_text(data = p_meds, aes(x = geno, y = med, label = med), 
              size = 3, vjust = -1.5)
```
> Q13: Read this file into R and determine the sample size for each genotype and their
corresponding median expression levels for each of these genotypes.

A:
```{r}
##Sample size of each genotype
table(expr$geno)
```

Their corresponding median expression levels are: A/A: 31.248, A/G: 25.065, G/G: 20.074.

> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative
expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3?

A: From this data, we see a trend where the presence of the G allele is linked with a decrease in gene expression, or the A allele is associated with an increase in gene expression. Individuals with the A/A genotype have the highest expression, and people with the A/G genotype have a medium level, and those with the G/G genotype have a relatively low expression.

This suggests that the SNP could indeed be affecting its expression. There is a potential dose-dependent relationship where two copies of the G allele (G/G genotype) leading to a reduction in expression compared to one copy (A/G genotype) or none (A/A genotype). Or it could be the vice versa, which an A allele leads to an increase in expression.
